---
title: "Data Exploration for Insight in the economy of an Massively Multiplayer Online Role-Playing Game"
author: "Lukas Zaiser"
params: 
 institute: "HdM Stuttgart"
 country: "Germany"
 year: "2022"
 data_date: "2022-04-01"
output:
 html_document: 
  css: style.css # define your own css
  df_print: paged #  tables are printed as HTML tables 
  highlight: default # syntax highlighting style 
  number_sections: yes # numbering of sections
  theme: paper # style option
  fig_height: 4 # figure height
  fig_width: 8 # figure width
  toc: yes # table of content
  toc_float: 
    collapsed: true # show full toc
    smooth_scroll: true # toc scrolling behavior
  includes:
    after_body: footer.html # include footer
---

```{r init, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(readr)

library(ggplot2)
library(gt)
library(gtable)
library(DT)

library(scales) # for label_number_si
```

# Load data

```{r load_csv, message=FALSE, warning=FALSE}
dir <- getwd()
date <- params$data_date

price_list_buys <- read.csv(paste(dir, "/gw2-all-buys-raw-", date, ".csv", sep = ""))
price_list_sells <- read.csv(paste(dir, "/gw2-all-sells-raw-", date, ".csv", sep = ""))
item_list <- read.csv(paste(dir, "/gw2-all-items-raw-", date, ".csv", sep = ""))

df_buys <- item_list %>% left_join(price_list_buys)
df_sells <- item_list %>% left_join(price_list_sells)

rm(dir)
rm(date)
rm(item_list)
rm(price_list_sells)
rm(price_list_buys)
```

## Data cleaning

```{r}
df_sells <- df_sells %>% 
  mutate(rarity = as.factor(rarity),
         type = as.factor(type),
         item_type = as.factor(item_type),
         item_weight_class = as.factor(item_weight_class),
         unit_price_gold = unit_price / 10000) %>% 
  select(-unit_price)

df_buys <- df_buys %>% 
  mutate(rarity = as.factor(rarity),
         type = as.factor(type),
         item_type = as.factor(item_type),
         item_weight_class = as.factor(item_weight_class),
         unit_price_gold = unit_price / 10000) %>% 
  select(-unit_price)
```


# Data Exploration

## Types

```{r type_distribution}
df_types <- df_sells %>% 
  distinct(type, name) %>%
  group_by(type) %>% 
  count(type)

df_types %>% 
  ggplot() +
  geom_bar(aes(x = type, y = n, fill = type), stat='identity') +
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position="bottom") +
  labs(title = "Type distribution", subtitle = "of all types", fill = "Types",
    x = "Type", y = "Count", caption = paste("Data from", params$data_date))
```

There are a lot of different types of items.
As we can see, Armor, Consumables and Weapons are the most common items.

This means there are very many different weapon, consumable and armor items in the game, but not that many different crafting materials and upgrade components which seems reasonable.

Armors and Weapons can be further broken down into sub types:

```{r item_type_distribution}
df_item_types <- df_sells %>% 
  filter(type %in% c("Armor", "Weapon")) %>% 
  distinct(type, item_type, item_weight_class, name) %>%
  group_by(type, item_type, item_weight_class) %>% 
  count(item_type)

df_item_types %>% 
  ggplot() +
  geom_bar(aes(x = item_type, y = n, fill = item_weight_class), stat='identity') +
  facet_grid(df_item_types$type) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3, hjust=1)) +
  labs(title = "Detailed item types", subtitle = "Split by armors and weapons", fill = "Item weight",
    x = "Item types", y = "Count", caption = paste("Data from", params$data_date))

rm(df_item_types)
```

For armors we have the distinction between **Light** (for mages), **Medium** (for rangers) and **Heavy** (for warriors). 
Clothing are special items for the style points.
There is no such distinction for weapons. Different classes can use different weapons, but all of them can be used by more than one class.


## Rarity of items

There are different rarities in Guild Wars that suggest how often this item will occur in the game:

```{r rarity_of_items, message=FALSE, warning=FALSE}
rarities <- c("Junk", "Basic", "Fine", "Masterwork", "Rare", "Exotic", "Ascended", "Legendary")
colors <- 
  c("Junk" = "#a3aeb9",
    "Basic" = "#000000",
    "Fine" = "#6a9cd9", 
    "Masterwork" = "#17911a", 
    "Rare" = "#f0cd25",
    "Exotic" = "#e9ad1a",
    "Ascended" = "#e64680",
    "Legendary" = "#53238a")

data.frame(rarities) %>% 
  ggplot() +
  geom_bar(aes(x = rarities, fill = rarities)) +
  scale_x_discrete(limits = rarities) +
  scale_fill_manual("Rarity", values = colors) + 
  theme(axis.text.y=element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

What is obvious by the names, rarity on the left is in general more common than rarity on the right.

One important information is that a lot of very rare items (`Ascended` and `Legendary`) can't be sold to other players. That's the reason we see much more `Rare` and `Exotic` items being sold for high prices than those really rare ones.
Junk is zero because it can only be sold at non-player vendors.

In our dataset, we have the following distribution:

```{r rarity_distribution}
df <- df_sells %>% 
  count(rarity) %>% 
  arrange(desc(n))

df %>% 
  ggplot() +
  geom_bar(aes(x = rarity, y = n, fill = rarity), stat='identity') +
  scale_x_discrete(limits = rarities) +
  scale_fill_manual("Rarity", values = colors) +
  labs(title = "Rarity distribution", subtitle = "of all raritires", fill = "Rarity",
    x = "Rarity", y = "Count", caption = paste("Data from", params$data_date))

relevant_rarities <- df %>% 
  filter(n > 100000)

relevant_rarities <- relevant_rarities$rarity
```


## Price distribution

Now let's not only look at the top five items, but at the distribution in general.

```{r rough_price_distribution}
df_sells_distribution <- df_sells %>% 
  group_by(name) %>% 
  summarise(mean_unit_price_gold = mean(unit_price_gold),
            type = unique(type),
            rarity = unique(rarity)) %>% 
  arrange(desc(mean_unit_price_gold))
  
df_sells_distribution %>% 
  ggplot() +
  geom_histogram(aes(x =  mean_unit_price_gold, fill = type), stat="count") +
  scale_x_binned(limits = c(0, 1000)) +
  labs(x = "Mean price", y = "Count",
       title = "Item price distribution", subtitle = "Items being sold (offers), in gold", 
       caption = paste("Data from", params$data_date))
```

As we can see, prices are common between 0 - 100 gold. What we also see is that there are a lot of different types of items.
We can strip this down so we only see the relevant types.

```{r}

relevant_types <- df_sells_distribution %>% 
  group_by(type) %>% 
  count() %>% 
  filter(n > 350)

df_sells_distribution <- df_sells_distribution %>% 
  filter(type %in% C(relevant_types$type))

df_sells_distribution %>% 
  ggplot() +
  geom_histogram(aes(x =  mean_unit_price_gold, fill = type), stat="count") +
  scale_x_binned(limits = c(0, 50), breaks = c(0, 10, 25, 50)) +
  facet_grid(. ~ type) + 
  labs(x = "Mean price", y = "Count",
       title = "Item price distribution", subtitle = "Items being sold (offers), in gold", 
       caption = paste("Data from", params$data_date))
```



## Crafting Material

Let's have a closer look on one of the most important resources in the game: Crafting material.

The base material can be collected everywhere in the world from trees, stones or herbs.
As a player, you can learn up to two professions that can improve those base materials to something better and useful.

Let's first check if there are any outliers in the data that we should clean.

```{r outliers_crafting}
df_crafting_sells <- df_sells %>% 
  filter(type == "CraftingMaterial")

df_crafting_buys <- df_buys %>% 
  filter(type == "CraftingMaterial")

ggplot() +  
  geom_boxplot(data = df_crafting_sells, aes(x = 'Sells', y = unit_price_gold)) +
  geom_boxplot(data = df_crafting_buys, aes(x = 'Buys', y = unit_price_gold)) +
  geom_hline(yintercept = 999, linetype="dashed", color = "red") +
  labs(title = "Outliers on crafting material", subtitle = "Baseline at 999 gold",
    x = "", y = "Price in gold", caption = paste("Data from", params$data_date))

```

We can see that on the buys, there aren't any outliers because the `human brain` is still working and no one buys overpriced goods.
On the other hand, on the sells there are some people who are exaggerating with the prices.
Let's strip everything at 1000 and above.

```{r overview_crafting, message=FALSE, warning=FALSE}
df_sells_crafting_material <-  df_sells %>% 
  filter(type == "CraftingMaterial",
         unit_price_gold < 1000) %>% 
  group_by(rarity, name) %>% 
   summarize(
    mean_price = mean(unit_price_gold),
    sum_quantity = sum(quantity),
    picture = web_image(unique(icon), height = 25)
  ) %>% 
  mutate(
    quantity_label = label_number_si(accuracy=0.1)(sum_quantity),
    mean_price_gold = round(mean_price, digits = 3)) %>% 
  arrange(desc(sum_quantity)) %>% 
  select(rarity, name, mean_price_gold, quantity_label, sum_quantity, picture)

df_sells_crafting_material %>% 
  datatable(extensions = c('ColReorder', 'Buttons', 'Responsive', 'Scroller', 'SearchPanes', 'Select'),
            options = list(colReorder = TRUE,
                           dom = 'Bfrtip', 
                           buttons = list('searchPanes', 'copy', list(
                                           extend = 'collection',
                                           buttons = c('csv', 'excel', 'pdf'), 
                                           text = 'Download'))),
            escape = FALSE,
            colnames=c("#", "Rarity", "Name", "Mean price in gold", "Quantity (short)", "Quantity", "Icon")
            )

exp_sell_item_name <- df_sells_crafting_material %>% 
  arrange(desc(mean_price_gold)) %>% 
  ungroup() %>%
  select(name) %>% 
  first() %>% 
  first()

```
